### 2019-01-16 Paper/main.tex
Updates from ShareLaTeX

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Shap-based Average Treatment Effect Calculations}
\author{Rossman, Hagai
        \and
        Yacovzada, Nancy-Sara
        \and
        Kalka, Iris}
\date{January 2019}

\begin{document}

\maketitle

\section{Abstract}

\section{Introduction}

\section{Methods}

\section{Results}

\section{Discussion}

\section{References}

\end{document}

### 2019-01-16 src/simulation_functions.py
Add files via upload

@@ -1,0 +1,41 @@
+import numpy as np
+
+
+def kang_shafer_data(random_seed=0):
+    """
+        Generate Y_true from Kang-Schafer Data Generating Process
+        
+        Parameters
+        ----------
+        random_seed : numpy random seed. default=0
+        
+        Inside
+        ------
+        Xs are auxiliary random variables, true_propensity, A is treatment variable
+        
+        Returns
+        -------
+        A: binary treatment assignment
+        Y_true: Simulated Y
+    
+        References
+        ----------
+        Joseph DY Kang, Joseph L Schafer, et al. Demystifying double robustness: A comparison of alternative strategies
+        for estimating a population mean from incomplete data. Statistical science, 22(4):523–539, 2007.
+        
+    """
+    np.random.seed = random_seed
+    X1 = np.random.normal(0,1)
+    X2 = np.random.normal(0,1)
+    X3 = np.random.normal(0,1)
+    X4 = np.random.normal(0,1)
+
+    true_propensity = 1/(1+np.exp(X1-0.5*X2+0.25*X3+0.1*X4))
+    A = int(true_propensity>=0.5)
+    # print(X1, X2, X3, X4, true_propensity, A)
+    
+    E_Y = 210 + A + 27.4*X1 + 13.78*X2 + 13.7*X3 + 13.7*X4
+
+    Y_true = np.random.normal(E_Y,1)
+    
+    return A, Y_true

### 2019-01-16 src/simulation_functions.py
Add files via upload

@@ -1,13 +1,15 @@
 import numpy as np
 
 
-def kang_shafer_data(random_seed=0):
+def kang_shafer_data(size=(1,) ,random_seed=0, X_observed=True):
     """
         Generate Y_true from Kang-Schafer Data Generating Process
         
         Parameters
         ----------
+        size : size of output arrays
         random_seed : numpy random seed. default=0
+        X_observed : Specify whether Xs are "not observed" , but a nonlinear transformation of them is
         
         Inside
         ------
@@ -25,17 +27,24 @@ def kang_shafer_data(random_seed=0):
         
     """
     np.random.seed = random_seed
-    X1 = np.random.normal(0,1)
-    X2 = np.random.normal(0,1)
-    X3 = np.random.normal(0,1)
-    X4 = np.random.normal(0,1)
+    X1 = np.random.normal(0,1,size=size)
+    X2 = np.random.normal(0,1,size=size)
+    X3 = np.random.normal(0,1,size=size)
+    X4 = np.random.normal(0,1,size=size)
 
     true_propensity = 1/(1+np.exp(X1-0.5*X2+0.25*X3+0.1*X4))
-    A = int(true_propensity>=0.5)
+    A = (true_propensity>=0.5).astype(int)
     # print(X1, X2, X3, X4, true_propensity, A)
     
     E_Y = 210 + A + 27.4*X1 + 13.78*X2 + 13.7*X3 + 13.7*X4
 
     Y_true = np.random.normal(E_Y,1)
     
-    return A, Y_true
+    if X_observed:
+        return X1, X2, X3, X4, A, Y_true
+    else: # Xs are "not observed" , but a nonlinear transformation of them is
+        X1_t = np.exp(X1/2)
+        X2_t = (X2/(1+np.exp(X1))) + 10
+        X3_t = (((X1*X3)/25) + 0.6)**3
+        X4_t = (X2 + X4 + 20)**2
+        return X1_t, X2_t, X3_t, X4_t, A, Y_true

### 2019-02-23 Paper/main.tex
I am the PI Iris
The draft I have

@@
         \and
-        Yacovzada, Nancy-Sara
+        Yacovzada, Nancy
         \and
@@
 \section{Introduction}
-
+Propensity score (PS) techniques, as formalized by Rosenbaum and Rubin \cite{rosenbaum1983central}, enable  controlling confounding in non experimental studies in medicine and epidemiology. 
+A central issue facing researchers using PS methods is how to select the variables to be included in the PS model.
+The bias and variance of the estimated exposure effect can depend strongly on which of these candidate variables are included in the PS model \cite{brookhart2006variable}.
+A possible limitation of PS methods, as described by Rubin in \cite{rubin1997estimating}, is that a covariate related to treatment assignment but unrelated to the outcome is handled the same as a covariate related to treatment but also strongly related to outcome.
+This can be a limitation of propensity scores, since inclusion of irrelevant covariates reduces the efficiency of an estimated exposure effect.
+However, it is claimed that if such a variable had even a weak effect on the outcome, the bias resulting from its exclusion would dominate any loss of efficiency for a reasonable-sized study.
 \section{Methods}
-
+\subsection{Data Simulations}
+The Z-Bias generation process \cite{myers2011effects}
 \section{Results}
@@
 
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
+
 \end{document}

### 2019-03-06 src/simulation_functions.py
Move treatment computing functions to .py

@@ -42,21 +42,21 @@ def kang_shafer_data(size=(1,), random_seed=0, X_observed=True):
     X4 = np.random.normal(0, 1, size=size)
 
     true_propensity = 1 / (1 + np.exp(X1 - 0.5 * X2 + 0.25 * X3 + 0.1 * X4))
-    A = (true_propensity >= 0.5).astype(int)
+    t = (true_propensity >= 0.5).astype(int)
     # print(X1, X2, X3, X4, true_propensity, A)
 
-    E_Y = 210 + A + 27.4 * X1 + 13.78 * X2 + 13.7 * X3 + 13.7 * X4
+    E_Y = 210 + t + 27.4 * X1 + 13.78 * X2 + 13.7 * X3 + 13.7 * X4
 
-    Y_true = np.random.normal(E_Y, 1)
+    y = np.random.normal(E_Y, 1)
 
     if X_observed:
-        return X1, X2, X3, X4, A, Y_true
+        return X1, X2, X3, X4, t, y
     else:  # Xs are "not observed" , but a nonlinear transformation of them is
         X1_t = np.exp(X1 / 2)
         X2_t = (X2 / (1 + np.exp(X1))) + 10
         X3_t = (((X1 * X3) / 25) + 0.6) ** 3
         X4_t = (X2 + X4 + 20) ** 2
-        return X1_t, X2_t, X3_t, X4_t, A, Y_true
+        return X1_t, X2_t, X3_t, X4_t, t, y
 
 
 def get_ACIC_2019_datasets(high_dim=True, test=True):

### 2019-03-06 src/auxiliary_methods.py
Creating one awful script with all our necessary functions... Good luck to us :-)

@@ -1,0 +1,246 @@
+from heapq import nsmallest
+
+import numpy as np
+import scipy.spatial.distance as distance
+import shap
+import sklearn
+
+
+def calc_distance(x, func="euclidean"):
+    """
+    Computes distances
+    :param x: Matrix of NxM where N is number of samples and M is number of features.
+    :param func: Distance metric: euclidean, mahalanobis
+    :return: Distance metric of NxN
+    """
+    if x.shape[1] == 1:
+        x = x.reshape(x.shape[0], 1)
+    return distance.squareform(distance.pdist(x, metric=func))
+
+
+def get_trained_model(x, y, base_estimator=sklearn.ensemble.gradient_boosting):
+    """
+    Trains the model and return it
+    :param x: matrix
+    :param y: predicted
+    :return: trained model
+    """
+    m = base_estimator()
+    m.fit(x, y)
+    return m
+
+
+def get_propensity_score(x, y):
+    """
+
+    :param x:
+    :param y:
+    :return:
+    """
+    m = get_trained_model(x, y)
+    return m.predict_proba(x)
+
+
+def get_inverse_propensity_weighing(x, y):
+    """
+
+    :param x:
+    :param y:
+    :return:
+    """
+    propensity = get_propensity_score(x, y)
+
+
+def calc_inv_propensity_score_weighting(df):
+    """
+    Inverse Propensity Score Weighting (IPW) for Average Treatment Affect for Treated (ATT)
+    Aim: to weight the treated and controlled observations to make them representative of the population of interest.
+
+    ATT = E[Y1−Y0|T=1] = E_X (E(Y1−Y0|X,T=1)) = E_X ( E(Y1|X,T=1) − E(Y0|X,T=1) )
+
+    Under SITA: E(Y0|X,T=1) = E(Y0|X,T=0)
+
+    Therefore: ATT = E_X ( E(Y1|X,T=1) − E(Y0|X,T=0) )
+
+    Using propensity score e(X) we obtain:
+    ATT = E_e(X) ( E(Y1|e(X),T=1) − E(Y0|e(X),T=0) )
+
+    The propensity score and inverse probability of treatment weighting in average treatment effect in the treated (ATT)
+    See:
+    https://onlinelibrary.wiley.com/doi/full/10.1002/sim
+    ##IPW_ATT = 1/n * (A - (np.dot(B, C)).sum()).6607
+    https://www.ncbi.nlm.nih.gov/pubmed/28436047
[…]

### 2019-03-07 src/simulation_functions.py
Fixing output of Kang-Safer simulation. And correcting random seed.

@@ -3,6 +3,7 @@ import sys
 from os import path
 
 import numpy as np
+import pandas as pd
 
 try:
     from src.simulation_iterator import SimulationIterator
@@ -35,7 +36,7 @@ def kang_shafer_data(size=(1,), random_seed=0, X_observed=True):
         for estimating a population mean from incomplete data. Statistical science, 22(4):523–539, 2007.
         
     """
-    np.random.seed = random_seed
+    np.random.seed(random_seed)
     X1 = np.random.normal(0, 1, size=size)
     X2 = np.random.normal(0, 1, size=size)
     X3 = np.random.normal(0, 1, size=size)
@@ -50,13 +51,13 @@ def kang_shafer_data(size=(1,), random_seed=0, X_observed=True):
     y = np.random.normal(E_Y, 1)
 
     if X_observed:
-        return X1, X2, X3, X4, t, y
+        return pd.DataFrame({'x1': X1, 'x2': X2, 'x3': X3, 'x4': X4, 'T': t, 'Y': y})
     else:  # Xs are "not observed" , but a nonlinear transformation of them is
         X1_t = np.exp(X1 / 2)
         X2_t = (X2 / (1 + np.exp(X1))) + 10
         X3_t = (((X1 * X3) / 25) + 0.6) ** 3
         X4_t = (X2 + X4 + 20) ** 2
-        return X1_t, X2_t, X3_t, X4_t, t, y
+        return pd.DataFrame({'x1_t': X1_t, 'x2_t': X2_t, 'x3_t': X3_t, 'x4_t': X4_t, 'T': t, 'Y': y})
 
 
 def get_ACIC_2019_datasets(high_dim=True, test=True):

### 2019-03-07 src/auxiliary_methods.py
Fixing bugs (now that I have finally run the code)

@@ -1,3 +1,4 @@
+import pandas as pd
 from heapq import nsmallest
 
 import numpy as np
@@ -18,14 +19,17 @@ def calc_distance(x, func="euclidean"):
     return distance.squareform(distance.pdist(x, metric=func))
 
 
-def get_trained_model(x, y, base_estimator=sklearn.ensemble.gradient_boosting):
+def get_trained_model(x, y, base_estimator='sklearn.ensemble.gradient_boosting.GradientBoostingRegressor'):
     """
     Trains the model and return it
     :param x: matrix
     :param y: predicted
+    :param base_estimator: String with the name of the base estimator
+        (default value is 'sklearn.ensemble.gradient_boosting.GradientBoostingRegressor')
     :return: trained model
     """
-    m = base_estimator()
+    __import__(base_estimator.split('.')[0])
+    m = eval(base_estimator + '()')
     m.fit(x, y)
     return m
 
@@ -49,6 +53,7 @@ def get_inverse_propensity_weighing(x, y):
     :return:
     """
     propensity = get_propensity_score(x, y)
+    return propensity
 
 
 def calc_inv_propensity_score_weighting(df):
@@ -99,24 +104,6 @@ def calc_inv_propensity_score_weighting(df):
     return IPW_ATT[0]
 
 
-# def get_matching_pairs(df, dist_func, n_neighbors=1):
-#     """
-#
-#     :param df:
-#     :param dist_func:
-#     :return:
-#     """
-#     df_0 = df.loc[df['T'] == 0].drop('T', axis=1)
-#     x_0 = df_0.drop('Y', axis=1)
-#     y_0 = df_0['Y']
-#     df_1 = df.loc[df['T'] == 1].drop('T', axis=1)
-#     x_1 = df_1.drop('Y', axis=1)
-#     y_1 = df_1['Y']
-#
-#     knn = sklearn.neighbors.KNeighborsClassifier(n_neighbors=n_neighbors, metric=dist_func, algorithm='ball_tree')
-#
-
-
 def calc_matching(df, dist_func='euclidean', k=1):
     """
     Computes the Average Treatment Effect on the treated through matching.
@@ -129,26 +116,26 @@ def calc_matching(df, dist_func='euclidean', k=1):
     :param k: Number of neighbors to consider.
     :return: An Average Treatment Effect scalar.
     """
-    x = df.loc[df['T'] == 0].drop(['T', 'Y'], axis=1)
[…]

### 2019-03-07 src/auxiliary_methods.py
Fixing more bugs and running Kang Shafer data.

@@ -34,14 +34,14 @@ def get_trained_model(x, y, base_estimator='sklearn.ensemble.gradient_boosting.G
     return m
 
 
-def get_propensity_score(x, y):
+def get_propensity_score(x, t):
     """
 
     :param x:
-    :param y:
+    :param t:
     :return:
     """
-    m = get_trained_model(x, y)
+    m = get_trained_model(x, t, 'sklearn.ensemble.gradient_boosting.GradientBoostingClassifier')
     return m.predict_proba(x)
 
 
@@ -93,10 +93,10 @@ def calc_inv_propensity_score_weighting(df):
     # List of outcomes Y for subjects that didn't receive treatment (T=0)
     B = df.loc[df['T'] == 0, ['Y']].values.reshape(-1)
 
-    propensity_scores = get_propensity_score(df.drop('Y', axis=1).df['Y'])
+    propensity_scores = get_propensity_score(df.drop(columns=['T', 'Y']), df['T'])
 
     # The weight for a control subject is defined as e(X)/1−e(X)
-    C = (propensity_scores / (1 - propensity_scores))[(df['T'] == 0).values].values
+    C = (propensity_scores / (1 - propensity_scores))[(df['T'] == 0).values]eht
 
     # The IPW estimator for ATT
     IPW_ATT = (1 / n1 * A - ((np.dot(B, C)).sum() / C.sum()))

### 2019-03-07 Paper/main.tex
Updates from Overleaf

--- 2019-01-16 Paper/main.tex
@@
 \title{Shap-based Average Treatment Effect Calculations}
-\author{Rossman, Hagai
+\author{Kalka, Iris
         \and
-        Yacovzada, Nancy-Sara
+        Yacovzada, Nancy
         \and
-        Kalka, Iris}
+        Rossman, Hagai}
 \date{January 2019}
@@
 
-\section{Abstract}
+\section{Introduction}
+Causal inference from observational data is a crucial an challenging field of work.
+Often works rely on assessing potential outcomes in order to estimate treatment effects \cite{neyman1923application}\cite{rubin1974estimating}.
+Many techniques, such as propensity score (PS) techniques, as formalized by Rosenbaum and Rubin \cite{rosenbaum1983central}, model observed data to predict treatment effects.
+These techniques enable  controlling confounding in non experimental studies in medicine and epidemiology.
 
-\section{Introduction}
+A central issue facing researchers using such methods is how to select the variables to be included in the model estimating for treatment effect.
+The bias and variance of the estimated exposure effect can depend strongly on which of these candidate variables are included in the PS model \cite{brookhart2006variable}.
+
+One assumption in the potential outcomes framework is ignorability, meaning that there are no unmeasured confounders \cite{rosenbaum1983central}.
+More specificly the assumption of ignorability demands that potential outcomes are independent of treatment assignment, conditioned on the set of observed covariates.
+
+A possible limitation of PS methods, as described by Rubin in \cite{rubin1997estimating}, is that a covariate related to treatment assignment but unrelated to the outcome is handled the same as a covariate related to treatment but also strongly related to outcome.
+Such inclusion of irrelevant covariates might reduce the efficiency of an estimated treatment effect \cite{10.1093/biomet/asx009}.
+In some cases the use of pretreatment covariates might even increase bias of treatment effect estimators.
+
+Our work is based on an assumption that every covariate can be treated as a covariate effecting both the treatment and the outcome.
+We believe that the size of the effect can be computed and taken into account such that pretreatment covariates are nearly ignored whereas other treatments are taken into account.
 
 \section{Methods}
+We compare treatment effect evaluation using several methods.
+We separate the methods to baseline methods already available in the literature: methods based on an estimator predicting the treatment; methods based on an estimator predicting the outcome; and methods based on two estimators one predicting treatment and the other predicting outcome.
 
-\section{Results}
+For all of these evaluations we used the Scikit-Learn gradient boosting predictors with default parameters \cite{scikit-learn}.
+
+\subsection{Baseline}
+
+The Z-Bias generation process \cite{myers2011effects}
+
+\subsection{Results}
 
@@
 
+\section{Future Work}
+
 \section{References}
 
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
+
 \end{document}

### 2019-03-10 Paper/main.tex
Methods

@@
 \usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 
+We compare methods by looking at the predicted Average Treatement Effect on Treated (ATT) compared with the true known ATT. Let us recall the definition of ATT:
+\begin{equation*}
+    ATT = \mathbb{E}[]Y_1 - Y_0 | T=1]
+\end{equation*}
+
 For all of these evaluations we used the Scikit-Learn gradient boosting predictors with default parameters \cite{scikit-learn}.
@@
 \subsection{Baseline}
+Our first method of estimating ATT is Inverse Propensity Score Weighting (IPW).
+We use the following equation for calculation of ATT:
+\begin{equation*}
+    \begin{split}
+        ATT = & \mathbb{E}[Y_1 - Y_0 | T=1] \\
+        & \mathbb{E}_X[\mathbb{E}[Y_1 - Y_0 | X, T=1]] \\
+        & \mathbb{E}_X[\mathbb{E}[Y_1 | X, T=1] - \mathbb{E}[Y_0 | X, T=1]]
+    \end{split}
+\end{equation*}
+
+Under strongly ignorable treatment assumption the potential outcomes $Y_0$, $Y_1$ are independent of the treatment selection given the observed covariates $X$. Therefore:
+\begin{equation*}
+    ATT = \mathbb{E}_X[\mathbb{E}[Y_1|X, T=1] - \mathbb{E}[Y_0 | X, T=0]]
+\end{equation*}
+
+Thus we use the propensity score and the inverse probability of treatment weighting to compute ATT \cite{abdia2017propensity}.
+
+The two other methods of baseline estimation relay on matching functions.
+For every treated individual we locate the most similar untreated individual.
+Thus estimating the treatment effect for every treated individual.
+This method uses the Scikit-learn Nearest neighbors tool for identification of the most similar individual \cite{scikit-learn}.
+
+Given a distance matrix between all individuals ATT is computed through matching by calculating the average difference in outcome for every treated individual.
+This is done by the difference between the individual's true outcome and the mean outcome of $k$ nearest untreated neighbors based on the given distance matrix.
+The total ATT is then the average of all treated individuals' average difference in outcome.
+
+Distance (and inversely similarity) between individuals is computed based solely on the propensity score computed for each individual.
+The two methods of distance used by us are Euclidean distance and Mahalanobis distance.
+
+\subsection{Based on treatment prediction}
+When accounting for the difference in effects of each covariate we chose to look at the feature importances in the treatment predictor.
+In order to make the different features importances comparable we use SHapley Additive exPlanations (SHAP) values \cite{lundberg2017unified}.
+These values are scalar values explaining the individual contribution of each covariate for each individual in the estimation of the treatment.
+
+Given a matrix of $N\times{}M$ covariates (where $N$ is the number of individuals and $M$ is the number of covariates) we create a single predictor $\mathcal{M}$ from the covariates to the treatment (marked a as a binary value).
+We then compute the SHAP values of $\mathcal{M}$ for every individual resulting in an $N\times{}M$ matrix.
+
+Our distances are computed based on the SHAp values matrix using either Euclidean of Mahalanobis distance for every two individuals. Afterwards the ATT is calculated using matching.
+
+\subsection{Based on outcome prediction}
+We create a single predictor that predicts the outcome given both an $N\times{}M$ matrix of covariates and a boolean vector of length $N$ of the treatment.
+After prediction of outcome we compute the SHAP values of all input values, giving us a matrix of size $N\times{}(M+1)$.
+
+Our computed distance should be reflective only on pretreatment features.
+It should not account for whether an individual is treated or not.
+Therefore, we remove the SHAP value respective of the treatment assignment, yielding us a matrix of $N\times{}M$.
+
+The given matrix is then used to compute distances either through Euclidean distance of Mahalanobis distance between every two individuals.
+ATT is then calculated using matching based on the distance measures.
+
+\subsection{Based on treatment prediction and outcome prediction}
+We create two predictors.
+One predictor $\mathcal{M}_T$ predicts the treatment assignment from covariates.The second predictor $mathcal{M}_Y$ predicts the outcome from both the covariates and the treatment assignment.
+
+We compute the SHAP values from both treatment and outcome predictions ($S_T$ and $S_Y$ respectively).
+In order to have comparable SHAP values we remove the column related to treatment from the $S_Y$ matrix.
+
+At this point each individual is represented with two vectors of length $M$ (number of covariates).
+We then join the vectors by dividing the values in $S_Y$ by those in $S_T$ in an element-wise division.
+Thus each person is represented with a single combined vector of length $M$.
+
+Distances between individuals are computed using either Euclidean or Mahalanobis distances on combined vectors. ATT is computed using matching based on the distance matrix.
+
+\subsection{Simulation}
 

### 2019-03-10 src/simulation_functions.py
Making treatment assignment random.

@@ -43,7 +43,7 @@ def kang_shafer_data(size=(1,), random_seed=0, X_observed=True):
     X4 = np.random.normal(0, 1, size=size)
 
     true_propensity = 1 / (1 + np.exp(X1 - 0.5 * X2 + 0.25 * X3 + 0.1 * X4))
-    t = (true_propensity >= 0.5).astype(int)
+    t = np.apply_along_axis(lambda p: np.random.binomial(1, p), 0, true_propensity)
     # print(X1, X2, X3, X4, true_propensity, A)
 
     E_Y = 210 + t + 27.4 * X1 + 13.78 * X2 + 13.7 * X3 + 13.7 * X4

### 2019-03-10 src/simulation_functions.py
Fix differences between paper and code.

@@ -46,7 +46,7 @@ def kang_shafer_data(size=(1,), random_seed=0, X_observed=True):
     t = np.apply_along_axis(lambda p: np.random.binomial(1, p), 0, true_propensity)
     # print(X1, X2, X3, X4, true_propensity, A)
 
-    E_Y = 210 + t + 27.4 * X1 + 13.78 * X2 + 13.7 * X3 + 13.7 * X4
+    E_Y = 210 + 27.4 * X1 + 13.78 * X2 + 13.7 * X3 + 13.7 * X4
 
     y = np.random.normal(E_Y, 1)
 

### 2019-03-10 src/auxiliary_methods.py
Nothing by a TYPO to see here.

@@ -65,7 +65,7 @@ def calc_inv_propensity_score_weighting(df):
 
     Under SITA: E(Y0|X,T=1) = E(Y0|X,T=0)
 
-    Therefore: ATT = E_X ( E(Y1|X,T=1) − E(Y0|X,T=0) )
+    Therefore: ATT = E11_X ( E(Y1|X,T=1) − E(Y0|X,T=0) )
 
     Using propensity score e(X) we obtain:
     ATT = E_e(X) ( E(Y1|e(X),T=1) − E(Y0|e(X),T=0) )
@@ -96,7 +96,7 @@ def calc_inv_propensity_score_weighting(df):
     propensity_scores = get_propensity_score(df.drop(columns=['T', 'Y']), df['T'])
 
     # The weight for a control subject is defined as e(X)/1−e(X)
-    C = (propensity_scores / (1 - propensity_scores))[(df['T'] == 0).values]eht
+    C = (propensity_scores / (1 - propensity_scores))[(df['T'] == 0).values]
 
     # The IPW estimator for ATT
     IPW_ATT = (1 / n1 * A - ((np.dot(B, C)).sum() / C.sum()))

### 2019-03-10 src/auxiliary_methods.py
Moving data_generation into source

@@ -161,12 +161,12 @@ def get_shap_values(model, x):
     return shap_values_df
 
 
-def calc_model_t_shap_matching(df, dist_func='euclidean', k=1):
+def calc_model_t_shap_matching(df, dist_func='euclidean', k=[1]):
     """
     Compute the Average Treatment Effect by matching based on Shap values of the model predicting T.
     :param df: A DataFrame of individuals.
     :param dist_func: A distance metric for Scipy's pdist function.
-    :param k: Number of neighbors to consider.
+    :param k: Number of neighbors to consider. This value comes as an array so that all possible values are considered.
     :return: Average Treatment Effect.
     """
     x = df.drop(columns=['Y', 'T'])
@@ -176,16 +176,19 @@ def calc_model_t_shap_matching(df, dist_func='euclidean', k=1):
     features_and_outcomes = shap_values.copy()
     features_and_outcomes['T'] = t
     features_and_outcomes['Y'] = df['Y']
-    ATT = calc_matching(features_and_outcomes, dist_func=dist_func, k=k)
-    return ATT
+    ATTs = []
+    for current_k in k:
+        ATT = calc_matching(features_and_outcomes, dist_func=dist_func, k=current_k)
+        ATTs.append(ATT)
+    return ATTs
 
 
-def calc_model_y_shap_matching(df, dist_func='euclidean', k=1):
+def calc_model_y_shap_matching(df, dist_func='euclidean', k=[1]):
     """
     Compute the Average Treatment Effect by matching based on Shap values of the model predicting Y.
     :param df: A DataFrame of individuals.
     :param dist_func: A distance metric for Scipy's pdist function.
-    :param k: Number of neighbors to consider.
+    :param k: Number of neighbors to consider. This value comes as an array so that all possible values are considered.
     :return: Average Treatment Effect.
     """
     x = df.drop(columns=['Y'])
@@ -201,18 +204,21 @@ def calc_model_y_shap_matching(df, dist_func='euclidean', k=1):
     features_and_outcomes.drop(columns=['T'], inplace=True)
     features_and_outcomes['T'] = t
     features_and_outcomes['Y'] = y
-    ATT = calc_matching(features_and_outcomes, dist_func=dist_func, k=k)
-    return ATT
+    ATTs = []
+    for current_k in k:
+        ATT = calc_matching(features_and_outcomes, dist_func=dist_func, k=current_k)
+        ATTs.append(ATT)
+    return ATTs
 
 
[…]

### 2019-03-10 src/data_generation.py

@@ -30,49 +30,41 @@
 
 ##################################
 ## Packages ##
-import logging
-from collections import defaultdict, Counter
 from datetime import datetime
 
-import matplotlib.pyplot as plt
-
 import numpy as np
-import pandas as pd
-import scipy.sparse as sp
-from sklearn.calibration import calibration_curve
-from sklearn.ensemble import RandomForestClassifier, GradientBoostingClassifier
-from sklearn.linear_model import LogisticRegression
-from sklearn.model_selection import StratifiedKFold, cross_val_predict, cross_val_score
 
-print(__doc__)
+# print(__doc__)
 
 
-class MonteCarloSimulation(object):
-
-    def __int__(self, n_simulations=2500, dataset_size=10000):
+class MonteCarloSimulation:
+    def __init__(self, n_simulations=2500, dataset_size=10000, random_seed=0):
 
         print("Start Monte Carlo Simulation ..", datetime.now())
 
         self.n_simulations = n_simulations
         self.dataset_size = dataset_size
+        self.random_seed = random_seed
 
-        #TODO: handle these params, we should play around with the values of the m
+        # TODO: handle these params, we should play around with the values of the m
         self.gamma0 = .3
-        self.gamma1 = .0 # 0.006, 0.06, 0.24, 0.6
+        self.gamma1 = .0  # 0.006, 0.06, 0.24, 0.6
         self.alpha0 = .3
-        self.alpha1 = .0 # 0.06, 0.18, 0.33
-        self.alpha2 = 0 # 0.06, 0.18, 0.33
-        self.beta0 = .2 #.01
-        self.beta1 = .0 # 0.08, 0.36, 0.5
-        self.beta2 = .0 # .2
+        self.alpha1 = .0  # 0.06, 0.18, 0.33
+        self.alpha2 = 0  # 0.06, 0.18, 0.33
+        self.beta0 = .2  # .01
+        self.beta1 = .0  # 0.08, 0.36, 0.5
+        self.beta2 = .0  # .2
+
 
         self.additive_data = self.sim_additive_model()
 
         self.multi_data = self.sim_multiplic_model()
 
-        self.save_data_to_file() # TODO: to implment
+        self.save_data_to_file()  # TODO: to implement
 
         print("Finish Monte Carlo Simulation ..", datetime.now())
+
 
     def sim_additive_model(self):
 
@@ -95,6 +87,7 @@ class MonteCarloSimulation(object):
         :return: data: numpy ndarray, matrix of matrices
         """
[…]

### 2019-03-10 Paper/main.tex
Updates from Overleaf

@@
         ATT = & \mathbb{E}[Y_1 - Y_0 | T=1] \\
-        & \mathbb{E}_X[\mathbb{E}[Y_1 - Y_0 | X, T=1]] \\
-        & \mathbb{E}_X[\mathbb{E}[Y_1 | X, T=1] - \mathbb{E}[Y_0 | X, T=1]]
+        = & \mathbb{E}_X[\mathbb{E}[Y_1 - Y_0 | X, T=1]] \\
+        = & \mathbb{E}_X[\mathbb{E}[Y_1 | X, T=1] - \mathbb{E}[Y_0 | X, T=1]]
     \end{split}
@@
 \subsection{Simulation}
+\subsubsection{Kang-Schafer}
+Our first method of simulation is the Kang-Schafer method for creating simulated data \cite{kang2007demystifying}.
+This method assigns four random covariates for every individual ($x_i^j~\mathcal{N}(0,1)$ for every $i\in{}\{1,2,3,4\}$).
+Propensity is then computed in the following way:
+\begin{equation*}
+    p(T=1| X=x) = \frac{1}{1 + \exp{(x_1 - \frac{x_2}{2} + \frac{x_3}{4} + \frac{x_4}{10})}}
+\end{equation*}
+
+We then assign treatment using the true propensity score as the probability for a binomial variable, yielding treatment assignments that are binary.
+
+Given both the covariates and the treatment assignments, we compute a expected outcome as:
+\begin{equation*}
+    \mathbb{E}[Y|X=x, T=t] = 210 + t + 27.4 \cdot{} x_1 + 13.78 \cdot{} x_2 + 13.7 \cdot{} x_3 + 13.7 \cdot{} x_4
+\end{equation*}
+
+Under the strong ignorability assumption and equations from the Kang-Schafer paper we can compute the true ATT using the following steps:
+\begin{equation*}
+    \begin{split}
+        ATT = & \mathbb{E}[Y_1 | X=x, T=1] - \mathbb{E}[Y_0 | X=x, T=1] \\
+        \mathbb{E}[Y_1|T=1, X=x] = & \mathbb{E}[Y|T=1, X=x] \\
+        \mathbb{E}[Y_0|T=1, X=x] = & \mathbb{E}[Y_0|T=0, X=x] = \mathbb{E}[Y|T=0, X=x] \\
+    \end{split}
+\end{equation*}
+Therefore $ATT=\mu{}^{(1)} - \mu{}^{(0)} = 200-220 = -20$ 
+%TODO: is this not actually 20?
+
+\subsubsection{Z-Bias}
 
@@
 
-\end{document}
+\end{document}Y1Y1

### 2019-03-11 src/auxiliary_methods.py
Adding a kang schafer image

@@ -1,10 +1,9 @@
-import pandas as pd
 from heapq import nsmallest
-
+import sklearn
 import numpy as np
+import pandas as pd
 import scipy.spatial.distance as distance
 import shap
-import sklearn
 
 
 def calc_distance(x, func="euclidean"):
@@ -104,7 +103,7 @@ def calc_inv_propensity_score_weighting(df):
     return IPW_ATT[0]
 
 
-def calc_matching(df, dist_func='euclidean', k=1):
+def calc_matching(df, dist_func='euclidean', k=[1], d=None):
     """
     Computes the Average Treatment Effect on the treated through matching.
     :param df: DataFrame with the following columns:
@@ -114,38 +113,42 @@ def calc_matching(df, dist_func='euclidean', k=1):
     :param dist_func: Distance metric for evaluating the distances between individuals.
         Will be used as a scipy.distance metric.
     :param k: Number of neighbors to consider.
+    :param d: A distance matrix, will be calculated if not given
     :return: An Average Treatment Effect scalar.
     """
     x = df.drop(['T', 'Y'], axis=1)
 
-    d = calc_distance(x.values, dist_func)
-
-    ATTs = []
+    if d is None:
+        d = calc_distance(x.values, dist_func)
 
     # Get the treated (T=1) and controls (T=0)
     idx_treated = [i for i, v in enumerate((df['T'] == 1).values) if v]
     idx_untreated = [i for i, v in enumerate((df['T'] == 0).values) if v]
 
+    ATTs = []
     # For every subject with T=1, calculate the average Y0 of its neighbors
-    for index in idx_treated:
-        # Get outcome of current treated
-        current_outcome = df.iloc[index]['Y']
+    for current_k in k:
+        current_ATTs = []
+        for index in idx_treated:
+            # Get outcome of current treated
+            current_outcome = df.iloc[index]['Y']
 
-        # Finding the k-closest neighbors in T=0 based on their propensity score
-        neighbors_indices = nsmallest(k, idx_untreated, key=lambda untreated_index: d[index, untreated_index])
+            # Finding the k-closest neighbors in T=0 based on their propensity score
+            neighbors_indices = nsmallest(current_k, idx_untreated,
+                                          key=lambda untreated_index: d[index, untreated_index])
 
[…]

### 2019-03-11 src/data_generation.py
Adding Nancy's changes

@@ -1,3 +1,4 @@
+
 """
     Data generation - Monte Carlo simulation Simulations
         - Additive model
@@ -30,41 +31,50 @@
 
 ##################################
 ## Packages ##
+import logging
+from collections import defaultdict, Counter
 from datetime import datetime
 
+import matplotlib.pyplot as plt
+
 import numpy as np
+import pandas as pd
+import scipy.sparse as sp
+from sklearn.calibration import calibration_curve
+from sklearn.ensemble import RandomForestClassifier, GradientBoostingClassifier
+from sklearn.linear_model import LogisticRegression
+from sklearn.model_selection import StratifiedKFold, cross_val_predict, cross_val_score
 
-# print(__doc__)
+print(__doc__)
 
 
-class MonteCarloSimulation:
-    def __init__(self, n_simulations=2500, dataset_size=10000, random_seed=0):
+class MonteCarloSimulation(object):
+
+    def __init__(self, n_simulations=1, dataset_size=10000, save_to_file=False):
 
         print("Start Monte Carlo Simulation ..", datetime.now())
 
         self.n_simulations = n_simulations
         self.dataset_size = dataset_size
-        self.random_seed = random_seed
 
-        # TODO: handle these params, we should play around with the values of the m
+        #TODO: handle these params, we should play around with the values of the m
         self.gamma0 = .3
-        self.gamma1 = .0  # 0.006, 0.06, 0.24, 0.6
+        self.gamma1 = .06 # 0.006, 0.24, 0.6
         self.alpha0 = .3
-        self.alpha1 = .0  # 0.06, 0.18, 0.33
-        self.alpha2 = 0  # 0.06, 0.18, 0.33
-        self.beta0 = .2  # .01
-        self.beta1 = .0  # 0.08, 0.36, 0.5
-        self.beta2 = .0  # .2
-
+        self.alpha1 = .18 # 0.06, 0.018, 0.33
+        self.alpha2 = 0 # 0.06, 0.18, 0.33
+        self.beta0 = .2 #.01
+        self.beta1 = .0 # 0.08, 0.36, 0.5
+        self.beta2 = .2 # .2
 
         self.additive_data = self.sim_additive_model()
 
         self.multi_data = self.sim_multiplic_model()
 
-        self.save_data_to_file()  # TODO: to implement
+        if save_to_file:
+            self.save_data_to_file() # TODO: to implment
 
[…]

### 2019-03-11 src/data_generation.py
Pycharm's reformat

@@ -1,4 +1,3 @@
-
 """
     Data generation - Monte Carlo simulation Simulations
         - Additive model
@@ -31,19 +30,10 @@
 
 ##################################
 ## Packages ##
-import logging
-from collections import defaultdict, Counter
 from datetime import datetime
-
-import matplotlib.pyplot as plt
 
 import numpy as np
 import pandas as pd
-import scipy.sparse as sp
-from sklearn.calibration import calibration_curve
-from sklearn.ensemble import RandomForestClassifier, GradientBoostingClassifier
-from sklearn.linear_model import LogisticRegression
-from sklearn.model_selection import StratifiedKFold, cross_val_predict, cross_val_score
 
 print(__doc__)
 
@@ -57,22 +47,22 @@ class MonteCarloSimulation(object):
         self.n_simulations = n_simulations
         self.dataset_size = dataset_size
 
-        #TODO: handle these params, we should play around with the values of the m
+        # TODO: handle these params, we should play around with the values of the m
         self.gamma0 = .3
-        self.gamma1 = .06 # 0.006, 0.24, 0.6
+        self.gamma1 = .06  # 0.006, 0.24, 0.6
         self.alpha0 = .3
-        self.alpha1 = .18 # 0.06, 0.018, 0.33
-        self.alpha2 = 0 # 0.06, 0.18, 0.33
-        self.beta0 = .2 #.01
-        self.beta1 = .0 # 0.08, 0.36, 0.5
-        self.beta2 = .2 # .2
+        self.alpha1 = .18  # 0.06, 0.018, 0.33
+        self.alpha2 = 0  # 0.06, 0.18, 0.33
+        self.beta0 = .2  # .01
+        self.beta1 = .0  # 0.08, 0.36, 0.5
+        self.beta2 = .2  # .2
 
         self.additive_data = self.sim_additive_model()
 
         self.multi_data = self.sim_multiplic_model()
 
         if save_to_file:
-            self.save_data_to_file() # TODO: to implment
+            self.save_data_to_file()  # TODO: to implment
 
         print("Finish Monte Carlo Simulation ..", datetime.now())
 
@@ -112,7 +102,7 @@ class MonteCarloSimulation(object):
             x = np.random.binomial(n=num_trials, p=self.alpha0 + self.alpha1 * u + self.alpha2 * z, size=n_obs)
             y = np.random.binomial(n=num_trials, p=self.beta0 + self.beta1 * u + self.beta2 * x, size=n_obs)
             # Merge z,x, and y into one matrix
[…]

### 2019-03-11 Paper/main.tex
Updates from Overleaf

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
 
@@
 Therefore $ATT=\mu{}^{(1)} - \mu{}^{(0)} = 200-220 = -20$ 
-%TODO: is this not actually 20?
+%TODO: is this actually -20?
 
 \subsubsection{Z-Bias}
+Our second method of simulation is based on adjusting for instrumental biases.
+We implemented two types of simulations according to the Z-Bias generation process as described in Myers \textit{et al} \cite{myers2011effects}.
 
-The Z-Bias generation process \cite{myers2011effects}
+The simulation creates examples where there are: one instrumental variable $Z$, one unobserved variable affecting both treatment and outcome $U$, a treatment $X$ and an outcome $Y$.
+All four variables are binary.
+These are all described in figure \ref{fig:z_bias_explanation} from the original paper.
+
+\begin{figure}
+    \centering
+    \includegraphics{Paper/images/z_bias_schema.png}
+    \caption{A figure from Myers \textit{et al} depicting the relationships between variables in simulated data.}
+    \label{fig:z_bias_explanation}
+\end{figure}
+
+There are two methods of simulation we use, an additive and a multiplicative version.
+In both versions the order of variables simulated is consistent in order to ensure that the risk of outcome would depend directly on $U$ and $X$ and indirectly on $Z$.
+
+In the additive version of the simulation the first variable to be simulated is $Z$ that is created with the probability $P(Z=1)=0.5$.
+Next the unobserved variable is simulated so that $P(U=1|Z)=\gamma{}_0 + \gamma{}_1\cdot{}Z$.
+Next the treatment is simulation using $P(X=1|U,Z) = \alpha{}_0 + \alpha{}_1 \cdot{} U + \alpha{}_2 \cdot{} Z$.
+Last we simulate the outcome using $P(Y=1|X,U) = \beta{}_0 + \beta{}_1 \cdot{} U + \beta{}_2 \cdot{} U$
 

### 2019-03-11 src/data_generation.py
Fixing disagreements between documentation and code

@@ -78,7 +78,7 @@ class MonteCarloSimulation(object):
             U, X and Y such that
                 P(U=1|Z) = gamma0 + gamma1 * Z
                 P(X=1|U,Z) = alpha0 + alpha1 * U + alpha2 * Z
-                P(Y=1|X,U) = beta0 + beta1 * X + beta2 * U
+                P(Y=1|X,U) = beta0 + beta1 * U + beta2 * X
 
         Notes:
             1. Order of variables simulated is important so that the risk of outcome would depend directly on U and X and indirectly on Z.
@@ -127,7 +127,7 @@ class MonteCarloSimulation(object):
             U, X and Y such that
                 P(U=1|Z) = gamma0 * (gamma1 ^ Z)
                 P(X=1|U,Z) = alpha0 * (alpha1 ^ U) * (alpha2 ^ Z)
-                P(Y=1|X,U) = beta0 * (beta1 ^ X) * (beta2 ^ U)
+                P(Y=1|X,U) = beta0 * (beta1 ^ U) * (beta2 ^ X)
 
         :return: data: numpy ndarray, matrix of matrices
         """

### 2019-03-11 Paper/main.tex

@@
     \centering
-    \includegraphics{Paper/images/z_bias_schema.png}
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_schema_complex.png}
     \caption{A figure from Myers \textit{et al} depicting the relationships between variables in simulated data.}
@@
 Next the treatment is simulation using $P(X=1|U,Z) = \alpha{}_0 + \alpha{}_1 \cdot{} U + \alpha{}_2 \cdot{} Z$.
-Last we simulate the outcome using $P(Y=1|X,U) = \beta{}_0 + \beta{}_1 \cdot{} U + \beta{}_2 \cdot{} U$
+Last we simulate the outcome using $P(Y=1|X,U) = \beta{}_0 + \beta{}_1 \cdot{} U + \beta{}_2 \cdot{} U$.
+
+In the multiplicative version of the simulation the order of simulation is identical to that in the additive version.
+Probabilities are computed in a different format as is explained in the following equations:
+$$P(Z=1)=0.5$$
+$$P(U=1|Z)=\gamma_0\cdot{}(\gamma_1^Z)$$
+$$P(X=1|U,Z)= \alpha_0\cdot(\alpha_1^U)\cdot(\alpha_2^Z)$$
+$$P(Y=1|X,U) = \beta_0 \cdot (\beta_1^U) \cdot (\beta_2^X)$$
 
 \subsection{Results}
+We simulated a data of $10,000$ individuals using the Kang-Schafer method explained above. We then use the forementioned methods to estimate the ATT. Results are depicted in
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/kang_schafer_ATE_estimations_by_neighbor.jpg}
+    \caption{Caption}
+    \label{fig:my_label}
+\end{figure}
 

### 2019-03-11 Paper/main.tex
Done for the day

@@
 All four variables are binary.
-These are all described in figure \ref{fig:z_bias_explanation} from the original paper.
+These are all described in Figure \ref{fig:z_bias_explanation} from the original paper.
 
@@
 \subsection{Results}
-We simulated a data of $10,000$ individuals using the Kang-Schafer method explained above. We then use the forementioned methods to estimate the ATT. Results are depicted in
+% TODO: Consider adding the figure of Y and T shap values in the Kang-Schafer simulation. This needs to be recreated using the current simulation but is still doable.
+
+We simulated a data of $10,000$ individuals using the Kang-Schafer method explained above.
+We then use the aforementioned methods to estimate the ATT.
+For each method we estimated the ATT values using matching on $k$ neighbors, where $k=1,2,...,30$.
+Results of depicted in Figure \ref{fig:kang_schafer}.
 
@@
     \includegraphics[width=\textwidth]{Paper/images/kang_schafer_ATE_estimations_by_neighbor.jpg}
-    \caption{Caption}
-    \label{fig:my_label}
+    \caption{\textbf{ATT Estimations for Kang-Schafer Simulation}: This image shows estimated values of ATT on the Kang-Schafer simulated data.
+True ATT value is $-20$ and is show in a black solid line.
+The inverse propensity weighing is marked in a purple solid line.
+The rest of the estimators are drawn in style according to the distance function used for nearest neighbor matches - Euclidean or Mahalanobis.
+Color of estimations are saperated by the values used in distance measurements of samples: baseline, based on treatment predicting model, based on outcome predicting model, based on the combination of outcome and treatment prediction models.
+The $x$ axis is the number of neighbors considered in the nearest neighbor algorithm.}
+    \label{fig:kang_schafer}
 \end{figure}
 
+We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods.
+In edition we can see that ATT estimations based solely on SHAP values from the model predicting treatment yielded the worst results.
+Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
+The best results we given with using SHAP values combination from a model predicting the treatment and a model predicting the outcome.
+
+For all models it is shown that both the Euclidean and the Mahalanobis distance functions yielded similar results, thus showing a minimal effect.
+
+In addition we can see that although there is some effect on the number of neighbors considered during matching, this effect occurs mostly for very small number of neighbors and is vastly smaller when taking into account more than five neighbors.
+
+\textbf{TODO: Add Z bias results here. }
+
 \section{Discussion}
+In the framework of potential outcomes an assumption of ignorability is usually made.
+This assumption relies on independence between treatment and outcome given covariates.
+However, reality shows that not all covariates should be considered, as some are in fact instrumental variables effecting only the treatment assignment directly.
+
+Our work aims to create a framework allowing consideration of all covariates while accounting for a difference in effect on treatment and outcome.
+
+Our results show the the shared consideration of both a model predicting the treatment and a model predicting the outcome can indeed overcome issues of instrumental variables.
+This added value is shown only be relevant when considering both model, where using each model independently has comparable results to baseline estimators.
+
+In addition our work shows little to no impact of the chosen distance function in case of matching when comparing Euclidean and Mahalanobis distances.
+We remain confident however that matching function based on domain knowledge will give a better estimation of treatment effects.
 
 \section{Future Work}
+The ease of use in our work is meant for cases of high-dimensional data as well as simple data.
+We believe that the framework created will work well on data with many covariates, however have yet to test it.
+
+Another type of data the framework should be tested on is data with more complex connections.
+Our current simulations are all based on rather simple relations between variables.
+A comparison to cases where there are covariate effecting treatment and/or outcome indirectly could give a better understanding of the framework's abilities.
+This is especially true since the entire framework can also be used on more complex predictors.
+
+Our current framework has only one method of combining the SHAP values of a treatment predicting model and an outcome predicting model.
+The rational behind using the ratio of SHAP values between the two models is to reduce the impact of covariates that effect the treatment assignment but not the outcome.
+Although fitting said goal, we hope to try additional methods of combining SHAP values.
+Specifically we considered using a logarithm of the ration, a ratio of logarithms, an additive factor of the values, etc.
+
+Another field we have yet to explore is the use of the SHAP values in order to perform feature selection on the data.
+Cases where there are many covariates can at times result in over-fitted models and thus biased predictions of treatment effect.
+We hope that accounting for the shared effect of a covariate both on the treatment and the outcome could allow to remove uneccessary covariates in a through a causal look.
 

### 2019-03-12 Paper/main.tex

@@
 
-\title{Shap-based Average Treatment Effect Calculations}
+\title{SHAP-based Average Treatment Effect Calculations}
 \author{Kalka, Iris
         \and
-        Yacovzada, Nancy
-        \and
-        Rossman, Hagai}
-\date{January 2019}
+        Yacovzada, Nancy}
+
+\date{March 2019}
 
@@
 Often works rely on assessing potential outcomes in order to estimate treatment effects \cite{neyman1923application}\cite{rubin1974estimating}.
-Many techniques, such as propensity score (PS) techniques, as formalized by Rosenbaum and Rubin \cite{rosenbaum1983central}, model observed data to predict treatment effects.
-These techniques enable  controlling confounding in non experimental studies in medicine and epidemiology.
+Many techniques, such as propensity score (PS) techniques, as formalized by Rosenbaum and Rubin \cite{rosenbaum1983central}, model observed data in order to predict treatment effects.
+These techniques enable controlling confounding in non-experimental studies in medicine and epidemiology.
 
-A central issue facing researchers using such methods is how to select the variables to be included in the model estimating for treatment effect.
-The bias and variance of the estimated exposure effect can depend strongly on which of these candidate variables are included in the PS model \cite{brookhart2006variable}.
+%TODO: ref 1, is that really what Iris meant to cite? "neyman1923application"
+
+A central issue researchers are facing using such methods is how to select the set of variables to be included in the model estimating for treatment effect.
+The bias and variance of the estimated treatment effect can depend strongly on which of these candidate variables are included in the PS model \cite{brookhart2006variable}.
 
 One assumption in the potential outcomes framework is ignorability, meaning that there are no unmeasured confounders \cite{rosenbaum1983central}.
-More specificly the assumption of ignorability demands that potential outcomes are independent of treatment assignment, conditioned on the set of observed covariates.
+More specifically, the assumption of ignorability demands that potential outcomes are independent of treatment assignment, conditioned on the set of observed covariates.
 
@@
 Such inclusion of irrelevant covariates might reduce the efficiency of an estimated treatment effect \cite{10.1093/biomet/asx009}.
-In some cases the use of pretreatment covariates might even increase bias of treatment effect estimators.
+In some cases, the use of pretreatment covariates might even increase bias of treatment effect estimators.
 
 Our work is based on an assumption that every covariate can be treated as a covariate effecting both the treatment and the outcome. 
+%TODO: REFINE
 We believe that the size of the effect can be computed and taken into account such that pretreatment covariates are nearly ignored whereas other treatments are taken into account.

### 2019-03-13 Paper/main.tex
Updates from Overleaf

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
-\title{SHAP-based Average Treatment Effect Calculations}
+\title{Estimating SHAP-based Treatment Effects}
 \author{Kalka, Iris
@@
 Causal inference from observational data is a crucial an challenging field of work.
-Often works rely on assessing potential outcomes in order to estimate treatment effects \cite{neyman1923application}\cite{rubin1974estimating}.
+Often works rely on assessing potential outcomes in order to estimate treatment effects \cite{rubin1974estimating}.
 Many techniques, such as propensity score (PS) techniques, as formalized by Rosenbaum and Rubin \cite{rosenbaum1983central}, model observed data in order to predict treatment effects.
@@
 
+% FROM THE MEIR PAOER In studies of exposure effect, measured and unmeasured factors that are associated with both exposure and outcome may confound the targeted causal effect.
+Estimating the exposure effect conditional on all confounding factors yields consistent estimates (1–4), so choosing which variables to use for adjustment in studies with many measured covariates is an important step for ensuring the validity of effect estimates.
+
 \section{Methods}
@@
 \begin{equation*}
-    ATT = \mathbb{E}[]Y_1 - Y_0 | T=1]
+    ATT = \mathbb{E}[Y_1 - Y_0 | T=1]
 \end{equation*}
@@
 \subsection{Baseline}
+\subsubsection{Inverse Propensity Score Weighting}
 Our first method of estimating ATT is Inverse Propensity Score Weighting (IPW).
@@
 
-The two other methods of baseline estimation relay on matching functions.
-For every treated individual we locate the most similar untreated individual.
-Thus estimating the treatment effect for every treated individual.
-This method uses the Scikit-learn Nearest neighbors tool for identification of the most similar individual \cite{scikit-learn}.
-
-Given a distance matrix between all individuals ATT is computed through matching by calculating the average difference in outcome for every treated individual.
-This is done by the difference between the individual's true outcome and the mean outcome of $k$ nearest untreated neighbors based on the given distance matrix.
-The total ATT is then the average of all treated individuals' average difference in outcome.
-
-Distance (and inversely similarity) between individuals is computed based solely on the propensity score computed for each individual.
-The two methods of distance used by us are Euclidean distance and Mahalanobis distance.
+\subsubsection{Matching}
+We utilize an additional method which serves as baseline estimation that relies on matching functions. Matching functions exploits known outcomes $Y_0$ in order to estimate the $T_0$ of the treated.
+For every treated individual $(T=1)$ in the cohort, we calculate his potential outcome $\bar{Y_0}$  as follows: 
+\begin{itemize}
+    \item Find his most similar $k$ untreated individuals $(T=0)$ 
+    \item Calculate $\bar{Y_0}$ - the mean outcome of the $k$ closest neighbors 
+\end{itemize}
+Now, for every treated individual we calculate his treatment effect $\widehat{T}$ by subtracting the mean outcome $\bar{Y_0}$ with the individual's true outcome $Y_1$. 
+Thus, given a distance matrix between all individuals we use matching to approximate the treatment effect for every treated individual. 
+These estimates allow us to calculate the total ATT, as explained above, by average of $\widehat{T}$ of all treated individuals'. 
+Distance (and inversely, similarity) between individuals is computed based solely on the propensity score computed for each individual. 
+The distance function we use here is either the Euclidean distance or Mahalanobis distance.
+We use the Scikit-Learn K-Nearest Neighbors algorithm for identification of the most similar individuals \cite{scikit-learn}.
 
 \subsection{Based on treatment prediction}
-When accounting for the difference in effects of each covariate we chose to look at the feature importances in the treatment predictor.
-In order to make the different features importances comparable we use SHapley Additive exPlanations (SHAP) values \cite{lundberg2017unified}.
-These values are scalar values explaining the individual contribution of each covariate for each individual in the estimation of the treatment.
+Here we build a single "treatment predictor" $\mathcal{M_T}$: a classifier that uses the $N\times{}M$ matrix of covariates to predict the treatment. 
+In order to account for the difference in effects of each covariate we chose to look at the features importance obtained based on this model. 
+To make the different features importances comparable we use SHapley Additive exPlanations (SHAP) values \cite{lundberg2017unified}. 
+These values are scalar values, explaining the individual contribution of each covariate for each individual in the estimation of the treatment.
 
-Given a matrix of $N\times{}M$ covariates (where $N$ is the number of individuals and $M$ is the number of covariates) we create a single predictor $\mathcal{M}$ from the covariates to the treatment (marked a as a binary value).
-We then compute the SHAP values of $\mathcal{M}$ for every individual resulting in an $N\times{}M$ matrix.
+Given a matrix of $N\times{}M$ covariates (where $N$ is the number of individuals and $M$ is the number of covariates) we create a single predictor $\mathcal{M_T}$ from the covariates to the treatment (marked a as a binary value).
+We then compute the SHAP values of $\mathcal{M_T}$ for every individual resulting in an $N\times{}M$ matrix $S_T$.
 
-Our distances are computed based on the SHAp values matrix using either Euclidean of Mahalanobis distance for every two individuals. Afterwards the ATT is calculated using matching.
+The distances between individuals are computed based on the SHAP values matrix $S$, using either Euclidean of Mahalanobis distance for every two individuals. 
+Finally, the ATT is calculated using matching.
 
 \subsection{Based on outcome prediction}
-We create a single predictor that predicts the outcome given both an $N\times{}M$ matrix of covariates and a boolean vector of length $N$ of the treatment.
-After prediction of outcome we compute the SHAP values of all input values, giving us a matrix of size $N\times{}(M+1)$.
+Here we build a single "outcome predictor" $\mathcal{M_Y}$: a classifier that predicts the outcome given both $N\times{}M$ matrix of covariates and a boolean vector of length $N$ of the treatment. 
+Based on this prediction model we compute the SHAP values of all input values which yields a $N\times{}(M+1)$ matrix.
 
-Our computed distance should be reflective only on pretreatment features.
-It should not account for whether an individual is treated or not.
-Therefore, we remove the SHAP value respective of the treatment assignment, yielding us a matrix of $N\times{}M$.
+The distances between individuals are computed based on the SHAP values matrix $S$, using either Euclidean of Mahalanobis distance for every two individuals. 
+Our computed distance should be reflective only on pretreatment features, regardless if the individual is received treatment or not.
+Therefore, we remove the SHAP value respective of the treatment assignment, yielding us a matrix $S_Y$ of $N\times{}M$.
 
 The given matrix is then used to compute distances either through Euclidean distance of Mahalanobis distance between every two individuals. 
-ATT is then calculated using matching based on the distance measures.
+ATT is then calculated using matching based on the distance matrix.
 
@@
 We create two predictors. 
-One predictor $\mathcal{M}_T$ predicts the treatment assignment from covariates.The second predictor $mathcal{M}_Y$ predicts the outcome from both the covariates and the treatment assignment.
+One predictor $\mathcal{M_T}$ predicts the treatment assignment from covariates.
+The second predictor $mathcal{M}_Y$ predicts the outcome from both the covariates and the treatment assignment.
 
@@
 We then join the vectors by dividing the values in $S_Y$ by those in $S_T$ in an element-wise division.
-Thus each person is represented with a single combined vector of length $M$.
+Thus, each individual is represented with a single combined vector of length $M$.
 
@@
 \subsection{Simulation}
+
+\subsubsection{Z-Bias}
+Recent theoretical studies have shown that conditioning on an instrumental variable (IV), a variable that is associated with the treatment but not associated with outcome except through treatment, can increase both bias and variance of treatment effect estimates.
+We follow the work from Myers \textit{et al} \cite{myers2011effects} where they suggested the "Z-Bias simulations", designed to provide insight into the problem of conditioning on potential IVs and adjusting for instrumental biases.
+Their results indicated that effect estimates which are conditional on a perfect IV or near-IV may have larger bias and variance than the unconditional estimate.
+However, they showed that in most scenarios considered, the increases in error due to conditioning were small compared with the total estimation error.
+Hence, in these cases minimizing unmeasured confounding should be the priority when selecting variables for adjustment, even at the risk of conditioning on IVs.
+The simulation creates examples where there is: an instrumental variable $Z$, an unobserved variable affecting both treatment and outcome $U$, a treatment $X$ and an outcome $Y$.
+All four variables are assumed to be binary.
+The causal graph is described in Figure \ref{fig:z_bias_explanation} from the original paper.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_schema_complex.png}
+    \caption{A figure from Myers \textit{et al} depicting the relationships between variables in simulated data.}
+    \label{fig:z_bias_explanation}
+\end{figure}
+
+There are two methods of simulation we use, an additive and a multiplicative version.
+In both versions the order of variables simulated is consistent in order to ensure that the risk of outcome would depend directly on $U$ and $X$ and indirectly on $Z$.
+
+In the \textbf{additive} version of the simulation the first variable to be simulated is $Z$ that is created with the probability $P(Z=1)=0.5$.
+Next the unobserved variable is simulated so that $P(U=1|Z)=\gamma{}_0 + \gamma{}_1\cdot{}Z$.
+Next the treatment is simulation using $P(X=1|U,Z) = \alpha{}_0 + \alpha{}_1 \cdot{} U + \alpha{}_2 \cdot{} Z$.
+Last we simulate the outcome using $P(Y=1|X,U) = \beta{}_0 + \beta{}_1 \cdot{} U + \beta{}_2 \cdot{} U$.
+
+In the \textbf{multiplicative} version of the simulation the order of simulation is identical to that in the additive version.
+However, probabilities are computed in a different format as is explained in the following equations:
+$$P(Z=1)=0.5$$
+$$P(U=1|Z)=\gamma_0\cdot{}(\gamma_1^Z)$$
+$$P(X=1|U,Z)= \alpha_0\cdot(\alpha_1^U)\cdot(\alpha_2^Z)$$
+$$P(Y=1|X,U) = \beta_0 \cdot (\beta_1^U) \cdot (\beta_2^X)$$
+
 \subsubsection{Kang-Schafer}
@@
 
+\subsection{Results}
+We simulated data of 10,000 individuals using every time a different simulation approach as explained below. 
+We then use the aforementioned methods to estimate the ATT.
+For each method we estimated the ATT values using matching on $k$-neighbors, where $k = 1, 2, ..., 30$.
+
 \subsubsection{Z-Bias}
-Our second method of simulation is based on adjusting for instrumental biases.
-We implemented two types of simulations according to the Z-Bias generation process as described in Myers \textit{et al} \cite{myers2011effects}.
+We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods. In edition we can see
+that ATT estimations based solely on SHAP values from the model predicting
+treatment yielded the worst results. Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
+The best results we given with using SHAP values combination from a model
+predicting the treatment and a model predicting the outcome.
+For all models it is shown that both the Euclidean and the Mahalanobis
+distance functions yielded similar results, thus showing a minimal effect.
+In addition we can see that although there is some effect on the number of
+neighbors considered during matching, this effect occurs mostly for very small
+number of neighbors and is vastly smaller when taking into account more than
+five neighbors.
 
-The simulation creates examples where there are: one instrumental variable $Z$, one unobserved variable affecting both treatment and outcome $U$, a treatment $X$ and an outcome $Y$.
-All four variables are binary.
-These are all described in Figure \ref{fig:z_bias_explanation} from the original paper.
+\subsubsection{Kang-Schafer method}
+Results of depicted in Figure 2. We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods. In edition we can see
+that ATT estimations based solely on SHAP values from the model predicting
+treatment yielded the worst results. Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
+The best results we given with using SHAP values combination from a model
+predicting the treatment and a model predicting the outcome.
+For all models it is shown that both the Euclidean and the Mahalanobis
+distance functions yielded similar results, thus showing a minimal effect.
+In addition we can see that although there is some effect on the number of
+neighbors considered during matching, this effect occurs mostly for very small
+number of neighbors and is vastly smaller when taking into account more than
+five neighbors.
+TODO: Ad
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Paper/images/z_bias_schema_complex.png}
-    \caption{A figure from Myers \textit{et al} depicting the relationships between variables in simulated data.}
-    \label{fig:z_bias_explanation}
-\end{figure}
-
-There are two methods of simulation we use, an additive and a multiplicative version.
-In both versions the order of variables simulated is consistent in order to ensure that the risk of outcome would depend directly on $U$ and $X$ and indirectly on $Z$.
-
-In the additive version of the simulation the first variable to be simulated is $Z$ that is created with the probability $P(Z=1)=0.5$.
-Next the unobserved variable is simulated so that $P(U=1|Z)=\gamma{}_0 + \gamma{}_1\cdot{}Z$.
-Next the treatment is simulation using $P(X=1|U,Z) = \alpha{}_0 + \alpha{}_1 \cdot{} U + \alpha{}_2 \cdot{} Z$.
-Last we simulate the outcome using $P(Y=1|X,U) = \beta{}_0 + \beta{}_1 \cdot{} U + \beta{}_2 \cdot{} U$.
-
-In the multiplicative version of the simulation the order of simulation is identical to that in the additive version.
-Probabilities are computed in a different format as is explained in the following equations:
-$$P(Z=1)=0.5$$
-$$P(U=1|Z)=\gamma_0\cdot{}(\gamma_1^Z)$$
-$$P(X=1|U,Z)= \alpha_0\cdot(\alpha_1^U)\cdot(\alpha_2^Z)$$
-$$P(Y=1|X,U) = \beta_0 \cdot (\beta_1^U) \cdot (\beta_2^X)$$
-
-\subsection{Results}
 % TODO: Consider adding the figure of Y and T shap values in the Kang-Schafer simulation. This needs to be recreated using the current simulation but is still doable.
@@
 In addition we can see that although there is some effect on the number of neighbors considered during matching, this effect occurs mostly for very small number of neighbors and is vastly smaller when taking into account more than five neighbors.
-
-\textbf{TODO: Add Z bias results here. }
 
@@
 We hope that accounting for the shared effect of a covariate both on the treatment and the outcome could allow to remove uneccessary covariates in a through a causal look.
-
+\section{Code}
+Link to code in \href{https://github.com/kalkairis/causal_inference_2018}{GitHub}
 \section{References}

### 2019-03-13 src/zbias_data_generation.py
Updates from Overleaf

@@ -1,0 +1,178 @@
+"""
+    Data generation - Monte Carlo simulation Simulations
+        - Additive model
+        - Multiplicative model
+
+
+    Assumed Causal diagram:
+        U       -       An unmeasured confounder
+        Z       -       Instrumental variable
+        (X,Y)   -       The exposure-outcome pair
+
+    Graph structure:
+        U -> X (alpha1)
+        U -> Y (beta1)
+        X -> Y (beta2)
+        Z -> U (gamma1)
+        Z -> X (alpha2)
+
+    The true "exposure effect" and target of estimation is beta2.
+
+    For simplicity and to reflect a common study framework, all variables are binary.
+
+    The measured covariate Z may act as a confounder or as an instrumental variable for the exposure-outcome pair (X, Y).
+    Note that Z is not a perfect instrument because it is associated with the unmeasured confounder U through gamma1
+    However, by varying the value of gamma1, we can explore the impacts of conditioning on Z when it is a perfect instrument and when it is a near-instrument or confounder.
+    As shown by Pearl (24), bias amplification may result even when the conditioning variable is not a perfect instrument.
+    Consider relatively large values of gamma1 to compare the risks of adjusting for an IV with the benefits of adjusting for a real confounder.
+
+"""
+
+##################################
+## Packages ##
+from datetime import datetime
+
+import numpy as np
+import pandas as pd
+
+print(__doc__)
+
+
+class MonteCarloSimulation(object):
+
+    def __init__(self, n_simulations=1, dataset_size=10000, save_to_file=False, random_seed=1976):
+
+        print("Start Monte Carlo Simulation ..", datetime.now())
+
+        self.n_simulations = n_simulations
+        self.dataset_size = dataset_size
+        self.random_seed = random_seed
+
+        # TODO: handle these params, we should play around with the values of the m
+        self.gamma0 = .3
+        self.gamma1 = .06  # 0.006, 0.24, 0.6
+        self.alpha0 = .3
+        self.alpha1 = .18  # 0.06, 0.018, 0.33
+        self.alpha2 = 0  # 0.06, 0.18, 0.33
+        self.beta0 = .2  # .01
+        self.beta1 = .0  # 0.08, 0.36, 0.5
+        self.beta2 = .2  # .2
+
+        self.additive_data = self.sim_additive_model()
+
+        self.multi_data = self.sim_multiplic_model()
+
+        if save_to_file:
+            self.save_data_to_file()  # TODO: to implment
[…]

### 2019-03-13 Paper/main.tex
Final version for submission

@@
 \usepackage{hyperref}
+\usepackage{float}
 
@@
 \section{Introduction}
+
+%TODO: REFINE INTRODUCTION!
+
 Causal inference from observational data is a crucial an challenging field of work.
@@
 These techniques enable controlling confounding in non-experimental studies in medicine and epidemiology.
-
-%TODO: ref 1, is that really what Iris meant to cite? "neyman1923application"
 
@@
 Our work is based on an assumption that every covariate can be treated as a covariate effecting both the treatment and the outcome. 
-%TODO: REFINE
-We believe that the size of the effect can be computed and taken into account such that pretreatment covariates are nearly ignored whereas other treatments are taken into account.
-
-% FROM THE MEIR PAOER In studies of exposure effect, measured and unmeasured factors that are associated with both exposure and outcome may confound the targeted causal effect.
-Estimating the exposure effect conditional on all confounding factors yields consistent estimates (1–4), so choosing which variables to use for adjustment in studies with many measured covariates is an important step for ensuring the validity of effect estimates.
+In this work we wish to examine several approaches for estimating the Average Treatment Effect for Treated (ATT) in studies with observed and unobserved variables that are associated with both treatment and outcome.
+We believe that during the estimation of the potential outcomes those variables may confound the targeted causal effect.
+We propose a novel approach to incorporate features importance scores in the construction of the distance matrix in which is used in the matching operation.
+These scores are based on a method called  SHapley Additive exPlanations (SHAP), described in details in \cite{lundberg2017unified}.
+We compere eight different approaches for obtaining pair-wise distances and test them on three types of simulation data.
 
@@
 Recent theoretical studies have shown that conditioning on an instrumental variable (IV), a variable that is associated with the treatment but not associated with outcome except through treatment, can increase both bias and variance of treatment effect estimates.
-We follow the work from Myers \textit{et al} \cite{myers2011effects} where they suggested the "Z-Bias simulations", designed to provide insight into the problem of conditioning on potential IVs and adjusting for instrumental biases.
+We follow the work from Myers \textit{et al} \cite{myers2011effects} where they suggested the "Z-Bias simulations", Monte Carlo simulation designed to provide insight into the problem of conditioning on potential IVs and adjusting for instrumental biases.
 Their results indicated that effect estimates which are conditional on a perfect IV or near-IV may have larger bias and variance than the unconditional estimate.
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 $$P(Y=1|X,U) = \beta_0 \cdot (\beta_1^U) \cdot (\beta_2^X)$$
+
+The true "exposure effect" and target of estimation is $\beta_2$. 
+For simplicity and to reflect a common study framework, all variables are binary.
+The measured covariate $Z$ may act as a confounder or as an IV for the treatment-outcome pair $(X, Y)$.
+Note that $Z$ is not a perfect instrument because it is associated with the unobsereved confounder $U$ through $\gamma_1$. 
+However, by varying the value of $\gamma_1$, we can explore the impacts of conditioning on $Z$ when it is a perfect instrument and when it is a near-instrument or confounder.
+A relatively large values of $\gamma_1$ can be considered to compare the risks of adjusting for an IV with the benefits of adjusting for a real confounder.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsubsection{Z-Bias}
-We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods. In edition we can see
-that ATT estimations based solely on SHAP values from the model predicting
-treatment yielded the worst results. Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
+First, we simulated data from the causal diagram as presented in Figure \ref{fig:z_bias_explanation} with the following coefficients:
+\begin{itemize}
+    \item $\gamma_0 = .3$
+    \item $\gamma_1 = .06$
+    \item $\alpha_0 = .3$
+    \item $\alpha_1 = .18$
+    \item $\alpha_2 = .18$ 
+    \item $\beta_0 = .2$
+    \item $\beta_1 = .36$ 
+    \item $\beta_2 = .2$
+\end{itemize}
+
+In the case of \textbf{Z-bias additive ATT estimation}, we can see that the worst estimation was obtained by the estimations based on two classifiers, the outcome prediction $\mathcal{M_Y}$ and treatment prediction $\mathcal{M_T}$.
+The following estimations resulted identical estimation error, regardless of the distance function, and they also are the best estimations: Matching, SHAP-based  treatment prediction $\mathcal{M_T}$ and SHAP-based outcome prediction $\mathcal{M_Y}$.
+The results suggest that there is no improvement in using the SHAP-based distance matrix compered to standard matching with propensity score and in fact the IPW baseline method was the second based estimator.
+Full results can be found in the IPython notebook.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_additive_ATT_estimations_by_k_neighbors.png}
+    \caption{Z bias additive ATT estimations by number of k}
+    \label{fig:z_bias_additive}
+\end{figure}
+
+In the case of \textbf{Z-bias multiplicative ATT estimation}, we can see that the worst estimation was obtained by the IPW baseline method. 
+The following estimations resulted identical estimation error, regardless of the distance function, and they also are the best estimations: Matching, SHAP-based  treatment prediction $\mathcal{M_T}$ and SHAP-based outcome prediction $\mathcal{M_Y}$.
+The estimations based on two classifiers, the outcome prediction $\mathcal{M_Y}$ and treatment prediction $\mathcal{M_T}$, resulted with higher estimation error than these methods but lower than IPW. 
+The results suggest that there is no improvement in using the SHAP-based distance matrix compered to standard matching with propensity score. Full results can be found in the IPython notebook.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_multiplicative_ATT_estimations_by_k_neighbors.png}
+    \caption{Z bias multiplicative ATT estimations by number of k}
+    \label{fig:z_bias_multi}
+\end{figure}
+
+Second, we simulated data with the following simulation coefficients:
+\begin{itemize}
+    \item $\gamma_0 = .3$
+    \item $\gamma_1 = .06$
+    \item $\alpha_0 = .3$
+    \item $\alpha_1 = .18$
+    \item $\alpha_2 = 0$ 
+    \item $\beta_0 = .2$
+    \item $\beta_1 = .0$ 
+    \item $\beta_2 = .2$
+\end{itemize}
+Note that here we assigned $\beta_1=0$ and $\alpha_2 = 0$ which result that $Z$ is not a IV (we cancelled the arrow from $Z$ to $X$).
+In addition, under these coefficients the unobserved variable $U$ doesn't effect the outcome.
+Results for the additive data are presented in Figure \ref{fig:z_bias_add_no_alpha2}. 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_additive_ATT_estimations_by_k_neighbors_no_alpha2.png}
+    \caption{Z bias additive ATT estimations by number of k neighbors - No IV}
+    \label{fig:z_bias_add_no_alpha2}
+\end{figure}
+
+Results for the multiplicative data are presented in Figure \ref{fig:z_bias_multi_no_alpha2}.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{Paper/images/z_bias_multiplicative_ATT_estimations_by_k_neighbors_no_alpha2.png}
+    \caption{Z bias multiplicative ATT estimations by number of k - No IV}
+    \label{fig:z_bias_multi_no_alpha2}
+\end{figure}
+
+For all models it is shown that both the Euclidean and the Mahalanobis distance functions yielded similar results.
+
+% TODO: effect of the number of neighbors considered during matching
+%TODO: BIAS VARIANCE ANALYSIS!
+
+\subsubsection{Kang-Schafer method}
+Results of depicted in Figure \ref{fig:kang_schafer}. 
+We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods.
+In edition we can see that ATT estimations based solely on SHAP values from the model predicting treatment yielded the worst results.
+Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
 The best results we given with using SHAP values combination from a model
@@
 
-\subsubsection{Kang-Schafer method}
-Results of depicted in Figure 2. We can see that IPW estimation gives an opposite direction of error in estimation compared to all other ATT estimation methods. In edition we can see
-that ATT estimations based solely on SHAP values from the model predicting
-treatment yielded the worst results. Following are baseline models and models based solely on the SHAP values from the model prediction the outecome.
-The best results we given with using SHAP values combination from a model
-predicting the treatment and a model predicting the outcome.
-For all models it is shown that both the Euclidean and the Mahalanobis
-distance functions yielded similar results, thus showing a minimal effect.
-In addition we can see that although there is some effect on the number of
-neighbors considered during matching, this effect occurs mostly for very small
-number of neighbors and is vastly smaller when taking into account more than
-five neighbors.
-TODO: Ad
-
 % TODO: Consider adding the figure of Y and T shap values in the Kang-Schafer simulation. This needs to be recreated using the current simulation but is still doable.
@@
 Our work aims to create a framework allowing consideration of all covariates while accounting for a difference in effect on treatment and outcome.
+It proposes a method to improve propensity-based matching by integrating the SHAP algorithm, a method that gives importance score for every variable in the model for every individual in the cohort.
 
-Our results show the the shared consideration of both a model predicting the treatment and a model predicting the outcome can indeed overcome issues of instrumental variables.
+In the case of Z-bias MC simulation our results showed no benefit using the SHAP scores as inout to the distance metric.
+In addition, both the Euclidean and the Mahalanobis distances yielded the same estimation curves for every $k$.
+Surprisingly, the shared consideration of both a model predicting the treatment and a model predicting the outcome didn't overcome the issue of instrumental variables and hence didn't add value compared to using each model independently.
+
+However, for the Kang-Schafer simulation our results show the the shared consideration of both a model predicting the treatment and a model predicting the outcome can indeed overcome issues of instrumental variables.
 This added value is shown only be relevant when considering both model, where using each model independently has comparable results to baseline estimators.
@@
 \section{Code}
-Link to code in \href{https://github.com/kalkairis/causal_inference_2018}{GitHub}
+Link to Python code on \href{https://github.com/kalkairis/causal_inference_2018}{GitHub}
 \section{References}

### 2019-03-13 src/zbias_data_generation.py

@@ -53,9 +53,9 @@ class MonteCarloSimulation(object):
         self.gamma1 = .06  # 0.006, 0.24, 0.6
         self.alpha0 = .3
         self.alpha1 = .18  # 0.06, 0.018, 0.33
-        self.alpha2 = 0  # 0.06, 0.18, 0.33
+        self.alpha2 = .18  # 0.06, 0.18, 0.33
         self.beta0 = .2  # .01
-        self.beta1 = .0  # 0.08, 0.36, 0.5
+        self.beta1 = .36  # 0.08, 0.36, 0.5
         self.beta2 = .2  # .2
 
         self.additive_data = self.sim_additive_model()
